################################################################################
##  
##  This program analyses NDVI dynamics based on the GIMMS and MODIS datasets.
##  
################################################################################
##
##  Copyright (C) 2013 Thomas Nauss, Tim Appelhans
##
##  This program is free software: you can redistribute it and/or modify
##  it under the terms of the GNU General Public License as published by
##  the Free Software Foundation, either version 3 of the License, or
##  (at your option) any later version.
##
##  This program is distributed in the hope that it will be useful,
##  but WITHOUT ANY WARRANTY; without even the implied warranty of
##  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the
##  GNU General Public License for more details.
##
##  You should have received a copy of the GNU General Public License
##  along with this program.  If not, see <http://www.gnu.org/licenses/>.
##
##  Please send any comments, suggestions, criticism, or (for our sake) bug
##  reports to admin@environmentalinformatics-marburg.de
##
################################################################################

Evaluation of NDVI Dynamics at Mt. Kilimanjaro
================================================================================

```{r, message=FALSE, warning=FALSE, eval=TRUE, fig.width = 15, fig.height = 15, echo=FALSE}
### Staging area ###############################################################
# Load libraries (echo=FALSE,)
library(lattice)
library(Kendall)
library(mgcv)
library(gamair)
library(stats)
library(bfast)
library (ggplot2)
library (mapproj)
library (maps)
library (maptools)

# Script from http://www.student.unibe.ch/user/phaenggi/rscripts.html
source("D:/temp/kilimanjaro_ndvi_dynamics/scripts/ndvidyn/src/snht.R")

# Check operation system and set paths.
actsys <- Sys.info()['sysname']
path.gis.win <- "D:/temp/kilimanjaro_ndvi_dynamics/gis"
path.gis.lin <- "/media/mobile/kilimanjaro_ndvi_dynamics/gis"
path.ghcn.win <- "D:/temp/lcd_kilimanjaro/rainfall/ghcn" 
path.ghcn.lin <- "/media/mobile/lcd_kilimanjaro/rainfall/ghcn"
```


NDVI dynamics between 1982 and 2006
--------------------------------------------------------------------------------
To analyise the vegetation dynamics between 1982 and 2006, the AVHRR-based GIMMS dataset is used. The original GIMMS values have been extracted for 99 pixels centered arround Mt. Kilimanjaro using SAGA GIS.
The NOAA-AVHRR systems used within this study are
NOAA 7 for 01/1982 to 02/1985
NOAA 9 for 03/1985 to 10/1988
NOAA 11 for 11/1988 to 08/1994
NOAA 9 for 09/1994 to 01/1995
NOAA 14 for 02/1995 to 10/2000
NOAA 16 for 11/2000 to 12/2003
NOAA 17 for 01/2004 to 12/2006
```{r, message=FALSE, warning=FALSE, eval=TRUE,echo=FALSE}
# Read original gimms values for 1982 to 2006 for 99 pixels arround Kilimanjaro,
# set headers using PIDs and convert date infomration to date object.
if (actsys == "Windows") {
  wd <- path.gis.win
  setwd(wd)
} else {
  wd <- path.gis.lin
  setwd(wd)
}
gndvi.org <- read.csv("kilimanjaro_pixels_gimms_srtm_gimms.csv",
                      header=TRUE, sep = ",")
colnames(gndvi.org) <- c(colnames(gndvi.org[1]), colnames(gndvi.org[2]),  
                              paste("PID", 
                                    apply(gndvi.org[4,3:length(gndvi.org)], 2,
                                          toString),
                                    sep="")[1:(length(gndvi.org)-2)])
gndvi.org$DATE <- as.Date(strptime(paste(gndvi.org$DATE), "%Y-%m-%d"))

# Compute time series analysis and plot information for each of the 99 datasets. 
i <- 3            # for testing only
gndvi.org.sdat <- subset(gndvi.org, gndvi.org$DATE > 1900)
pdf("gndvi_org.pdf")
gndvi.org.sdat.mk11 <- list()
gndvi.org.sdat.bfast <- list()
for (j in 1:(length(gndvi.org)-2)) {
  i <- j+2
  print(j)
  print(gndvi.org[4,i])
  gndvi.org.sdat.ts <- ts(gndvi.org.sdat[, i], 
                         start=as.numeric(format(head(gndvi.org.sdat$DATE, 1), 
                                                 format = "%Y")), freq=12)

  # STL
  gndvi.org.sdat.ts.stl <- stl((gndvi.org.sdat.ts),
                               s.window="periodic", t.window=12)
  
  gndvi.org.sdat.ts.stl300 <- stl((gndvi.org.sdat.ts),
                               s.window="periodic", t.window=300)
  
  # 11-year running Mann-Kendall trend
  gndvi.org.sdat.mk11[[j]] <- sapply(seq(15), function(k) {
    date.start <- 1981 + k
    date.end <- date.start + 11
    gndvi.org.sdat.ts.stl.trend <- gndvi.org.sdat.ts.stl$time.series[,'trend']
    gndvi.org.sdat.ts.stl.trend.sub <- gndvi.org.sdat.ts.stl.trend[
      time(gndvi.org.sdat.ts.stl.trend)>= date.start &
      time(gndvi.org.sdat.ts.stl.trend)< date.end]
    temp.mk <- cor.test(gndvi.org.sdat.ts.stl.trend.sub,
                        time(seq(from=date.start, to=date.end-1/12, by=1/12)),
                        method="kendall", alternative = "two.sided")
    return.mk <- list(PID = j,
                      DATEMEAN = date.start + (date.end-date.start-1)/2,
                      tau = temp.mk$estimate,
                      p = temp.mk$p.value)
    return(return.mk)
  })
  
  # Correlation
  gndvi.org.sdat.sm <- cor.test(gndvi.org.sdat.ts.stl300$time.series[,'trend'],
                                time(gndvi.org.sdat$DATE),
                                method="spearman", alternative = "two.sided")
  gndvi.org.sdat.mk <- cor.test(gndvi.org.sdat.ts.stl300$time.series[,'trend'],
                                time(gndvi.org.sdat$DATE),
                                method="kendall", alternative = "two.sided")

  # BFAST
  rdist <- 10/length(gndvi.org.sdat.ts)
  gndvi.org.sdat.ts.bfit <- bfast(gndvi.org.sdat.ts, h=rdist,
                                  season="harmonic", max.iter=1, breaks=5)
  gndvi.org.sdat.ts.bfit.niter <- length(gndvi.org.sdat.ts.bfit$output)
  gndvi.org.sdat.ts.bfit.final <- gndvi.org.sdat.ts.bfit$output[[
    gndvi.org.sdat.ts.bfit.niter]]
  gndvi.org.sdat.ts.bfit.final.breakpoints <- round(as.numeric(
    format(head(gndvi.org.sdat$DATE, 1),format = "%Y")) + 
    gndvi.org.sdat.ts.bfit.final$bp.Vt$breakpoints/12, digits=0)

  plot(gndvi.org.sdat.ts.stl,
         main = paste("PID: ", gndvi.org[4,i],
                      ", MK: ", round(gndvi.org.sdat.mk$estimate, digits = 3),
                      ", SC:", round(gndvi.org.sdat.sm$estimate, digits = 3),
                      ", Years: ", toString(gndvi.org.sdat.ts.bfit.final.breakpoints),
                      sep=""))
  
  gndvi.org.sdat.bfast[[j]] <- c(PID = gndvi.org[4,i],
                            MK = round(gndvi.org.sdat.mk$estimate, digits = 3),
                            SM = round(gndvi.org.sdat.sm$estimate, digits = 3),
                            NIter = gndvi.org.sdat.ts.bfit.niter,
                            BP = gndvi.org.sdat.ts.bfit.final.breakpoints)
}
dev.off()
gndvi.org.sdat.mk11 <- data.frame(sapply(
  data.frame(t(do.call("cbind", gndvi.org.sdat.mk11))), unlist))
gndvi.org.sdat.bfast <- data.frame(t(do.call("cbind", gndvi.org.sdat.bfast)))
write.table(gndvi.org.sdat.mk11, file="test-mk.csv", sep = ";", dec = ",")
write.table(gndvi.org.sdat.bfast, file="test-bf.csv", sep = ";", dec = ",")

pdf("test.pdf")
graphics.off()
par(mfrow=c(10,11)) 
par(mar=c(1,1,1,4)*0.5) 
par(oma=c(1,1,1,1)*1.5) 
dummy <- sapply(seq(length(gndvi.org.sdat)-2), function(i){
  gndvi.org.sdat.mk11.subset <- subset(gndvi.org.sdat.mk11,
                                       gndvi.org.sdat.mk11$PID == i)
  gndvi.org.sdat.bfast.subset <- subset(gndvi.org.sdat.bfast,
                                        gndvi.org.sdat.bfast$PID == i)
#   plot(gndvi.org.sdat.mk11.subset$tau ~ gndvi.org.sdat.mk11.subset$DATE,
#        col = 'red', type = 'o',
#        ylim = range(-0.5, 0.5), xlab = "Date", ylab = "Mann-Kendall's tau",
#        main = paste("Mann-Kendalls's tau, ",
#                     "PID ", i, sep=""))
  plot(gndvi.org.sdat.mk11.subset$tau ~ gndvi.org.sdat.mk11.subset$DATE,
       col = 'red', type = 'o', xlab="", ylab="", main="", labels = FALSE,
       ylim = range(-0.5, 0.5))
  abline(h = 0, col = 'grey')
  abline(v = c(gndvi.org.sdat.bfast.subset
               [5:length(gndvi.org.sdat.bfast.subset)]),
         col = 'blue', lty = "dotted")
})
  
hist(gndvi.org.stat$BYEAR)
hist(gndvi.org.stat$PID, breaks=2)

subset(gndvi.org.stat, gndvi.org.stat$BYEAR > 1900)
gndvi.org.stat[with(gndvi.org.stat, order(BYEAR)), ]
plot(gndvi.org.stat$PID ~ gndvi.org.stat$BYEAR, labels=TRUE)
plot(gndvi.org.stat$PID ~ gndvi.org.stat$MK, labels=TRUE)
x <- 1:11
y <- 1:9
t <- data.frame(aggregate(gndvi.org.stat$MK ~ gndvi.org.stat$PID,FUN="mean")[,2])
t$PID <- seq(1:99)
t$lat <- gndvi.org[3,101]
t$lon <- gndvi.org[2,101]
t1 <- matrix(t, 11,9)
filled.contour(x,y,t1, axes=TRUE, )




# for (i in 3:length(gndvi.org)) {
#   print(gndvi.org[4,i])
#   gndvi.org.sdat.ts <- ts(gndvi.org.sdat[, i], 
#                          start=as.numeric(format(head(gndvi.org.sdat$DATE, 1), 
#                                                  format = "%Y")), freq=12)
# 
#   # Correlation
#   gndvi.org.sdat.sm <- cor.test(gndvi.org.sdat[, i], time(gndvi.org.sdat$DATE),
#                            method="spearman", alternative = "two.sided")
#   gndvi.org.sdat.mk <- cor.test(gndvi.org.sdat[, i], time(gndvi.org.sdat$DATE),
#                            method="kendall", alternative = "two.sided")
#   
# #   # Linear model
# #   gndvi.org.sdat.lm <- lm(gndvi.org.sdat[, i] ~ gndvi.org.sdat$DATE)
# 
#   # STL
# #   gndvi.org.sdat.ts.stl <- stl((gndvi.org.sdat.ts),
# #                                s.window="periodic", t.window=12)
# #   gndvi.org.sdat.ts.lm <- lm(gndvi.org.sdat.ts ~ time(gndvi.org.sdat.ts))
# #   gndvi.org.sdat.stl120 <- stl((gndvi.org.sdat.ts),
# #                                s.window="periodic", t.window=120)
# #   gndvi.org.sdat.stl300 <- stl((gndvi.org.sdat.ts),
# #                                s.window="periodic", t.window=300)
# 
#   # BFAST
#   rdist <- 10/length(gndvi.org.sdat.ts)
#   gndvi.org.sdat.ts.bfit <- bfast(gndvi.org.sdat.ts, h=rdist,
#                                   season="harmonic", max.iter=1, breaks=5)
#   gndvi.org.sdat.ts.bfit.niter <- length(gndvi.org.sdat.ts.bfit$output)
#   gndvi.org.sdat.ts.bfit.final <- gndvi.org.sdat.ts.bfit$output[[
#     gndvi.org.sdat.ts.bfit.niter]]
#   gndvi.org.sdat.ts.bfit.final.breakpoints <- round(as.numeric(
#     format(head(gndvi.org.sdat$DATE, 1),format = "%Y")) + 
#     gndvi.org.sdat.ts.bfit.final$bp.Vt$breakpoints/12, digits=0)
# 
# #   # GAM
# #   gndvi.org.sdat$MONTH <- as.numeric(format(gndvi.org.sdat$DATE, format = "%m"))
# #   gndvi.org.sdat$DATEN <- time(time(gndvi.org.sdat.ts))
# # 
# #   gndvi.org.sdat.mod <- gamm(gndvi.org.sdat[, i] ~ s(MONTH) + s(DATEN),
# #                         data = gndvi.org.sdat)
# #   gndvi.org.sdat.mod.pred <- predict(gndvi.org.sdat.mod$gam, 
# #                                 newdata = gndvi.org.sdat, type = "terms")
# #   gndvi.org.sdat.mod.pred.rains <- attr(gndvi.org.sdat.mod.pred, "constant") + 
# #                                    gndvi.org.sdat.mod.pred[,1]
# #   gndvi.org.sdat.mod.pred.raina <- attr(gndvi.org.sdat.mod.pred, "constant") + 
# #                                    gndvi.org.sdat.mod.pred[,2]
# 
#   # Plot STL result
#   print(paste("PID: ", gndvi.org[4,i],
#         ", MK: ", round(gndvi.org.sdat.mk$estimate, digits = 3),
#         ", SC: ", round(gndvi.org.sdat.sm$estimate, digits = 3),
#         ", Years: ", toString(gndvi.org.sdat.ts.bfit.final.breakpoints)))
# #   par(mfrow=c(1,1)) 
# #    plot(gndvi.org.sdat.ts.stl,
# #         main = paste("PID: ", gndvi.org[4,i],
# #                      ", MK: ", round(gndvi.org.sdat.mk$estimate, digits = 3),
# #                      ", SC:", round(gndvi.org.sdat.sm$estimate, digits = 3),
# #                      ", Years: ", toString(gndvi.org.sdat.ts.bfit.final.breakpoints),
# #                      sep=""))
# 
# #   # Plot BFAST results
# #   plot(gndvi.org.sdat.ts.bfit, main = i)
# #   plot(gndvi.org.sdat.ts.bfit, ANOVA=TRUE)
# #   plot(gndvi.org.sdat.ts.bfit, type="trend", largest=TRUE)
# #   plot(gndvi.org.sdat.ts.bfit, type="all")
# #   par(mfrow=c(2,1))
# #   plot(gndvi.org.sdat.ts.bfit.final$Vt, ylim = range(gndvi.org.sdat.ts.bfit.final$Vt))
# #   par(new=TRUE)
# #   plot(gndvi.org.sdat.ts.bfit.final$Tt, col='blue', ylim = range(gndvi.org.sdat.ts.bfit.final$Vt))
# #   str(gndvi.org.sdat.ts.bfit.final$bp.Vt$breakpoints)
# #   plot(gndvi.org.sdat.ts.bfit.final$Wt, ylim = range(gndvi.org.sdat.ts.bfit.final$Wt))
# #   par(new=TRUE)
# #   plot(gndvi.org.sdat.ts.bfit.final$St, col='red', ylim = range(gndvi.org.sdat.ts.bfit.final$Wt))
#   
#   # Plot combined STL results
# #   par(new=FALSE)
# #   plot(gndvi.org.sdat.ts, 
# #        main = paste(i, "MK-tau/MK-P/R2/R2", 
# #                     round(gndvi.org.sdat.mk$estimate, digits = 3), 
# #                     round(gndvi.org.sdat.mk$p.value, digits = 3), 
# #                     round(summary(gndvi.org.sdat.ts.lm)$r.squared, digits = 4),
# #                     round(summary(gndvi.org.sdat.lm)$r.squared, digits = 4),
# #                     sep=", "), 
# #        xlab = "Time", ylab = "NDVI",
# #        xlim = range(1982, 2006), ylim = range(gndvi.org.sdat.ts),
# #        axes=T, type = 'l', col = 'green')
# #   abline(gndvi.org.sdat.ts.lm, col = "darkgreen", lwd = 2)
# #   par(new=TRUE)
# #   plot(gndvi.org.sdat.stl$time.series[,'trend'], 
# #        xlab = "", ylab = "",
# #        xlim = range(1982, 2006), ylim = range(gndvi.org.sdat.ts),
# #        axes=T, type = 'l', col = 'red', lwd = 2)
# #   par(new=TRUE)
# #   plot(gndvi.org.sdat.stl120$time.series[,'trend'], 
# #        xlab = "", ylab = "",
# #        xlim = range(1982, 2006), ylim = range(gndvi.org.sdat.ts),
# #        axes=T, type = 'l', col = 'purple', lwd = 2)
# #   par(new=TRUE)
# #   plot(gndvi.org.sdat.stl300$time.series[,'trend'], 
# #        xlab = "", ylab = "",
# #        xlim = range(1982, 2006), ylim = range(gndvi.org.sdat.ts),
# #        axes=T, type = 'l', col = 'orange', lwd = 2)
# # #   par(new=TRUE)
# # #   plot(gndvi.elev.org.mod.pred.raina,
# # #        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
# # #        xlab = "", ylab = "", xaxt = "n",
# # #        axes=T, type = 'l', col = "red")
# #   par(new=TRUE)
# #   plot(act.station.stl$time.series[,'trend'], 
# #          xlab = "", ylab = "",
# #          xlim = range(1982, 2006),
# #          ylim = range(-500, 2500),
# #          axes=F, type = 'l', col = 'blue')
# #   axis(side=4, ylim = range(0, 2000))
# #   legend('top', c("GIMMS", "lm trend", "stl 18m", "stl 120m", "stl 300m",
# #                   paste("Rainfall", act.station_name, act.station_country, 
# #                         sep = " ")), 
# #          cex=0.5, col=c("green", "darkgreen", "red", "purple", "orange", "blue"), lty = 1)
# }











# gndvi.elev.org.levels <- colnames(gndvi.elev.org[3:length(gndvi.elev.org)])
# 
# # GHCN dataset for selected stations in Africa
# if (actsys == "Windows") {
#   wd <- path.ghcn.win
#   setwd(wd)
# } else {
#   wd <- path.ghcn.lin
#   setwd(wd)
# }
# ghcn <- read.csv("noaa-ncdc_64603_monthly_summaries_GHCN-Daily_east_africa.csv", 
#                  header=TRUE, sep = ";")
# ghcn$DATE <- as.Date(strptime(ghcn$DATE , "%Y%m%d"))
# ghcn$TPCP <- ghcn$TPCP / 10.0
# ghcn <- ghcn[order(ghcn$LATITUDE, decreasing = TRUE), ]
# 
# 
# # Analyse and plot dataset for each height level
# #par(mfrow=c(1,1))
# i <- "H3000"
# stations <- "GHCND:KE000063740"
# 
# for (i in gndvi.elev.org.levels) {
#   # Correlation
#   gndvi.elev.org.sm <- cor.test(gndvi.elev.org[, i], time(gndvi.elev.org$DATE), 
#                            method="spearman", alternative = "two.sided")
#   gndvi.elev.org.mk <- cor.test(gndvi.elev.org[, i], time(gndvi.elev.org$DATE), 
#                            method="kendall", alternative = "two.sided")
#   # Linear model
#   gndvi.elev.org.lm <- lm(gndvi.elev.org[, i] ~ gndvi.elev.org$DATE)
# 
#   # STL
#   gndvi.elev.org.ts <- ts(gndvi.elev.org[, i], 
#                          start=as.numeric(format(head(gndvi.elev.org$DATE, 1), 
#                                                  format = "%Y")), freq=12)
#   gndvi.elev.org.stl <- stl((gndvi.elev.org.ts),s.window="periodic", t.window=12)
#   gndvi.elev.org.ts.lm <- lm(gndvi.elev.org.ts ~ time(gndvi.elev.org.ts))
#   gndvi.elev.org.stl120 <- stl((gndvi.elev.org.ts),s.window="periodic", t.window=120)
#   gndvi.elev.org.stl300 <- stl((gndvi.elev.org.ts),s.window="periodic", t.window=300)
#   
#   # BFAST
#   rdist <- 10/length(gndvi.elev.org.ts)
#   gndvi.elev.org.bfit <- bfast(gndvi.elev.org.ts, h=rdist, season="harmonic", max.iter=1, breaks=5)
# 
#   # GAM
#   gndvi.elev.org$MONTH <- as.numeric(format(gndvi.elev.org$DATE, format = "%m"))
#   gndvi.elev.org$DATEN <- time(time(gndvi.elev.org.ts))
#   
#   
#   #gndvi.elev.org.mod <- gamm(gndvi.elev.org[, i] ~ s(MONTH, bs="cc") + s(DATEN, bs="cr"),
#   #                      data = gndvi.elev.org)
#   gndvi.elev.org.mod <- gamm(gndvi.elev.org[, i] ~ s(MONTH) + s(DATEN),
#                         data = gndvi.elev.org)
#   gndvi.elev.org.mod.pred <- predict(gndvi.elev.org.mod$gam, 
#                                 newdata = gndvi.elev.org, type = "terms")
#   gndvi.elev.org.mod.pred.rains <- attr(gndvi.elev.org.mod.pred, "constant") + 
#                                    gndvi.elev.org.mod.pred[,1]
#   gndvi.elev.org.mod.pred.raina <- attr(gndvi.elev.org.mod.pred, "constant") + 
#                                    gndvi.elev.org.mod.pred[,2]
# 
#   # GHCN Rainfall trend
#   act.station <- subset(ghcn, ghcn$STATION == stations)
#   act.station <- subset(act.station, 
#                         act.station$DATE == unique(act.station$DATE) &
#                         act.station$TPCP < 9000.0)
#   act_quantile <- quantile(act.station$TPCP, 0.95)
#   act.station <- subset(act.station, act.station$TPCP < act_quantile)
#   act.station_name <- unique(act.station$STATION_NAME)
#   act.station_country <- unique(act.station$COUNTRY)
#   act.station.ts <- ts(act.station$TPCP, 
#                        start=as.numeric(format(head(act.station$DATE, 1), format = "%Y")), freq=12)
#   act.station.stl <- stl((act.station.ts),s.window="periodic")
#   act.station.ts.lm <- lm(act.station.ts ~ time(act.station.ts))
# 
#   # Plot STL results
#   plot(gndvi.elev.org.stl, main = paste(i, "MK", 
#                                    round(gndvi.elev.org.mk$estimate, digits = 3),
#                                    sep=", "),)
# 
#   # Plot BFAST results
#   plot(gndvi.elev.org.bfit, main = i)
#   plot(gndvi.elev.org.bfit, ANOVA=TRUE)
#   plot(gndvi.elev.org.bfit, type="trend", largest=TRUE)
#   plot(gndvi.elev.org.bfit, type="all")
#   niter <- length(gndvi.elev.org.bfit$output)
#   out <- gndvi.elev.org.bfit$output[[niter]]
#   
#   # Plot combined STL results
#   par(new=FALSE)
#   plot(gndvi.elev.org.ts, 
#        main = paste(i, "MK-tau/MK-P/R2/R2", 
#                     round(gndvi.elev.org.mk$estimate, digits = 3), 
#                     round(gndvi.elev.org.mk$p.value, digits = 3), 
#                     round(summary(gndvi.elev.org.ts.lm)$r.squared, digits = 4),
#                     round(summary(gndvi.elev.org.lm)$r.squared, digits = 4),
#                     sep=", "), 
#        xlab = "Time", ylab = "NDVI",
#        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
#        axes=T, type = 'l', col = 'green')
#   abline(gndvi.elev.org.ts.lm, col = "darkgreen", lwd = 2)
#   par(new=TRUE)
#   plot(gndvi.elev.org.stl$time.series[,'trend'], 
#        xlab = "", ylab = "",
#        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
#        axes=T, type = 'l', col = 'red', lwd = 2)
#   par(new=TRUE)
#   plot(gndvi.elev.org.stl120$time.series[,'trend'], 
#        xlab = "", ylab = "",
#        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
#        axes=T, type = 'l', col = 'purple', lwd = 2)
#   par(new=TRUE)
#   plot(gndvi.elev.org.stl300$time.series[,'trend'], 
#        xlab = "", ylab = "",
#        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
#        axes=T, type = 'l', col = 'orange', lwd = 2)
# #   par(new=TRUE)
# #   plot(gndvi.elev.org.mod.pred.raina,
# #        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
# #        xlab = "", ylab = "", xaxt = "n",
# #        axes=T, type = 'l', col = "red")
#   par(new=TRUE)
#   plot(act.station.stl$time.series[,'trend'], 
#          xlab = "", ylab = "",
#          xlim = range(1982, 2006),
#          ylim = range(-500, 2500),
#          axes=F, type = 'l', col = 'blue')
#   axis(side=4, ylim = range(0, 2000))
#   legend('top', c("GIMMS", "lm trend", "stl 18m", "stl 120m", "stl 300m",
#                   paste("Rainfall", act.station_name, act.station_country, 
#                         sep = " ")), 
#          cex=0.5, col=c("green", "darkgreen", "red", "purple", "orange", "blue"), lty = 1)
#   
#   
# # Plot STL results overlayed in one plot and with linear model, GAMS prediction and rainfall
# #   par(new=FALSE)
# #   plot(gndvi.elev.org.ts, 
# #        main = paste(i, "MK-tau/MK-P/R2/R2", 
# #                     round(gndvi.elev.org.mk$estimate, digits = 3), 
# #                     round(gndvi.elev.org.mk$p.value, digits = 3), 
# #                     round(summary(gndvi.elev.org.ts.lm)$r.squared, digits = 4),
# #                     round(summary(gndvi.elev.org.lm)$r.squared, digits = 4),
# #                     sep=", "), 
# #        xlab = "Time", ylab = "NDVI",
# #        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
# #        axes=T, type = 'l', col = 'green')
# #   abline(gndvi.elev.org.ts.lm, col = "darkgreen")
# #   #axis(side=1, xlim = range(1982, 2006))
# #   #axis(side=2, ylim = range(gndvi.elev.org.ts))
# #   par(new=TRUE)
# #   plot(gndvi.elev.org.stl$time.series[,'trend'], 
# #        xlab = "", ylab = "",
# #        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
# #        axes=T, type = 'l', col = 'red')
# #   par(new=TRUE)
# #   plot(gndvi.elev.org.stl$time.series[,'seasonal'], 
# #        xlab = "", ylab = "",
# #        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
# #        axes=T, type = 'l', col = 'blue')
# #   par(new=TRUE)
# #   plot(gndvi.elev.org.mod.pred.raina,
# #        xlim = range(1982, 2006), ylim = range(gndvi.elev.org.ts),
# #        xlab = "", ylab = "", xaxt = "n",
# #        axes=T, type = 'l', col = "red")
# #   par(new=TRUE)
# #   plot(act.station.stl$time.series[,'trend'], 
# #          xlab = "", ylab = "",
# #          xlim = range(1982, 2006),
# #          ylim = range(-500, 2500),
# #          axes=F, type = 'l', col = 'blue')
# #   axis(side=4, ylim = range(0, 2000))
# #   legend('top', c("Satellite observation","Trend", "Linear", 
# #                   paste("Rainfall", act.station_name, act.station_country, 
# #                         sep = " ")), 
# #          cex=0.5, col=c("green", "red", "darkgreen", "blue"), lty = 1)
# #   
#   # Plot GAM results    
#   plot(gndvi.elev.org[, i] ~ DATE, data = gndvi.elev.org, 
#        main = paste(i, "MK-tau/MK-P/R2/R2", 
#                     round(gndvi.elev.org.mk$estimate, digits = 3), 
#                     round(gndvi.elev.org.mk$p.value, digits = 3), 
#                     round(summary(gndvi.elev.org.ts.lm)$r.squared, digits = 4),
#                     round(summary(gndvi.elev.org.lm)$r.squared, digits = 4),
#                     sep=", "), 
#        xlab = "Time", ylab = "NDVI",
#        col = "green", type = "l")
#   lines(gndvi.elev.org.mod.pred.raina ~ DATE, data = gndvi.elev.org, 
#         col = "red", lwd = 2)
#   lines(gndvi.elev.org.mod.pred.rains ~ DATE, data = gndvi.elev.org, 
#         col = "blue", lwd = 1, lty = 2)
#   abline(gndvi.elev.org.lm, col = "darkgreen")
#   legend('top', c("Satellite observation","Trend", "Seasonal", "Linear"), 
#          cex=0.5, col=c("green", "red", "blue", "darkgreen"), lty = 1)
# 
#   plot(gndvi.elev.org.mod$gam, pages = 1, 
#        main = paste(i, "MK", round(gndvi.elev.org.mk$estimate, digits = 3), 
#                     sep=", "))
# }
```

